---
title: "Examen R"
output: html_notebook
---
Instructions:
  
  1.- Create a numeric vector that contains the months of a year.
Then, create another vector containing the name for each month. Later, you should assign the months and names.
Finally, you should select and display the months related to the summer season (June, July, August)


```{r}
months <- c(1,2,3,4,5,6,7,8,9,10,11,12)
months_names <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sept", "Oct", "Nov", "Dec")
year <- matrix(months_names, months, byrow = TRUE)
year
summer <- year[6:8]
summer
```

2. Construct a matrix with 5 rows containing the numbers 1 up to 15, filled row-wise.

```{r}
nums <- c(1:15)
nums_mat <- matrix(nums, ncol = 5, byrow = TRUE)
nums_mat
```

3. Create a factor with these observations "XXL", "XXL", "XXL", "XL", "L", "M", "M", "S", "S", "S" and print a summary

```{r}
tams <- c("XXL", "XXL", "XXL", "XL", "L", "M", "M", "S", "S", "S")
factor_tams <- factor(tams)
factor_tams

sum_factor_tams <- summary(factor_tams)
sum_factor_tams
```

4.- Create and print a data frame that should contain:

5 of your courses (e.g. Fundamentals of Programming, Object-Oriented Programming, Databases, etc ).
5 of your marks
5 of your professors' names

```{r}
courses <- c("Inovacion Teconologica", "Topicos Selectos", "Programacion", "Base de Datos", "Ingenieria del Conocimiento")
marks <- c(8.2, 7.3, 9.1, 8.9, 7)
profs <- c("Abraham Lopez", "Alan Ponce", "Pancho Lopez", "Beatriz Gomez", "Vicente Fernandez")
semester <- data.frame(courses, marks, profs)
semester
```

5.- Load the dataset labeled as ITU_WB.csv that is located in the data folder
and storage this dataset in a variable called itu.
Then, you should describe the composition of the dataset.
This description must contain the number of observations, variables and the type of each variable

```{r}
itu <- read.csv("~/UACJ/10mo SEM/Int a la Ing del Conocimiento/ts-2019-project-Jeisiva/data/ITU_WB.csv")
ncol(itu)
nrow(itu)
total <- ncol(itu)*nrow(itu)
total
names_col <- names(itu)
names_col
lapply(itu, class)
glimpse(itu)
#En este DataSet podemos encontrar 12 columnas y 3735 lineas, dandonos un total de 44820 datos.
#Las variables (columnas) son:
# "num","Economy","Code","Region","Income.group","year","Internet.value","Total.population","GDP..US..","Life.Expectancy..Years.","GDP.per.capita.income","Fertility..Births.per.woman."
#El tipo de variable que tiene cada columna es:
    #num = "integer"
    #Economy = "factor"
    #Code = "factor"
    #Region = "factor"
    #Income.group = "factor"
    #year = "integer"
    #Internet.value = "numeric"
    #Total.population = "integer"
    #GDP..US.. = "numeric"
    #Life.Expectancy..Years. = "numeric"
    #GDP.per.capita.income = "numeric"
    #Fertility..Births.per.woman. = "numeric"

```

6.- Filter the itu dataset to retrieve only the observation from Mexico, grouped by year and in descending order selecting the "Internet value" column

```{r}
itu %>% filter(Economy == Mexico) %>% arrange(desc(Internet.Value)) %>% group_by(year)
```

7.- Create a new variable called itu_by_year grouping per "year" and "Income group" variables. Then, remove NAs observations. Later, using the function summarize() you should estimate the median of the Internet value variable and store it in a variable called "medianValue"

```{r}
itu %>% filter(Economy == Mexico) %>% arrange(desc(Internet.Value)) %>% group_by(year)
itu_by_year <- group_by(year, income.group)
na.omit(itu_by_year)

summarize <- function(){
  medianValue <- median(Internet.Value)
  medianValue
}
```

8.-Create a line plot to visualize trends over time. You should use the variables created in the itu_by_year dataset

```{r}

ggplot(itu_by_year, aes(x = year, y= income.group)) + geom_line()

```


9.- Create a new variable called itu_latin filtering per Region, selecting Latin America & Caribbean and grouping per Income group.

Then, using this dataset, create a plot visualizing the relationship between Internet value and GDP (US$) colored per Income group and faceting per Economy

```{r}
itu_latin <- itu %>% filter(Region == "Latin America", Region =="Caribbean") %>% group_by(Income.group)
ggplot(itu_latin, aes(x = GDP..US.., y = Internet.value, color = Income.group)) + geom_point() + face_wrap(~ Economy)

```

10.- Create a new variable called itu_2017 filtering the year 2017 and removing observations that contain NAs.

Then, using this dataset, create a plot visualizing the relationship between Internet value and GDP (US$), colored per Income group. Furthermore, you should use a log scale in both axes ("x" and "y"). Finally, you should use the facet function in order to wrap by Region.

```{r}
itu_2017 <- itu %<% na.omit(filter(year == 2017))
ggplot(itu_2017, aes(x = Internet.value, y = GDP..US.., color = Income.group)) + scale_x_log10() + scale_y_log10() + face_wrap(~ Region)
```


